MLH1 Results Outline

### 2020-02-19 doc/Results.rmd
added rec.group col to main bivData DF

load(file="~./MLH1repo/data/CleanBivData_13.2.20.RData")

@@

@@

@@

#code intp chr

@@

@@

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)
```

@@

@@

# Q2 SC

@@

Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Dom2

Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Musc2

Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Mol

@@

**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.

The positive correlation of sc lengths has a positive correlation with the number of COs on single chromosomes ,

in the musculus and molossinus strains, the positive correlation is more

In Domisticus the -- postive correlation only holds for the first 3 classes of bivalents (0CO, 1CO, 2CO) -- this illustrates how the differences between proportions of chromosome classes across strains confonds comparisons of SC length across strains.

The glm's for the full dataset across -- the coefficients are listed below

```{r wacky, echo=FALSE}

real.long.bivData$bin <- "long"
real.short.bivData$bin <- "short"

lng.shrt <- rbind(real.long.bivData[,c(1:2, 17,19)], real.short.bivData[,c(1:2, 17,19)])

uu <- ggplot(lng.shrt, aes(x=fileName, y=chromosomeLength) )  + geom_point()+facet_wrap(~strain, scales="free")

@@

## Total SC

@@

```{r show.tot.sc, echo=FALSE}

mouse.av_sc_male <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")
mouse.av_sc_male

@@

```{r Q2.SC_class.rec.group_setup, echo=FALSE}

@@

```{r Q2.SC.Length.stuff, echo=FALSE}
#all pooled SC lengths
#WSB     G   LEW  MOLF   KAZ CZECH 
#    5     9     5     4     3     3

mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
mouse.av.long.biv.table <- add_mouse(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)

mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)

mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )

mouse.av.short.biv.table <- add_mouse(mouse.av.short.biv.table)
mouse.av.short.biv.table <- add_strain(mouse.av.short.biv.table)
mouse.av.short.biv.table <- add_subsp(mouse.av.short.biv.table)
mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
```

```{r Q2.SC.long.biv_2, echo=FALSE}
shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")

```{r Q2.SC.ploy.male, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect

```{r Q2.glm.SC.long, echo=FALSE}
glm.Q2.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)

table(mouse.av.long.biv.table$mouse)

glmQ2.M2.SC.male_long.biv <- glm(mean.SC ~ strain, data= mouse.av.long.biv.table)
round(summary(glmQ2.M2.SC.male_long.biv)$coefficients[,c(1,4)], 6)
```
In the cell average Long biv data set, MSM G and CZECH strains have sig strain effects

very few strains have significant values for the mouse

```{r Q2.glm.SC.short, echo=FALSE}
glmQ2_SC_mouse.av.short_biv <- glm(mean.SC ~ strain, data= mouse.av.short.biv.table)

round(summary(glmQ2_SC_mouse.av.short_biv)$coefficients[,c(1,4)], 6)
```
Basically all strains have a significant effect for the mouse.av of long biv

```{r Q2.glm.SC.all.strain, echo=FALSE}

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)

@@

I think these comparisons should be done with the **Bivalent observations**
The logistic regression showed rec groups could be predicted by SC length - re-run them while separating out by chrm class

```{r Q2.SC.Chrm.Class, echo=FALSE}

@@

```{r Q2.SC.log.regression.show.plots, echo=FALSE}

@@

```{r Q2.SClength.ttest, echo=FALSE}
#MOUSE AVE, all
Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

@@

The top t.tests indicate that when all the mice averages are pooled, there's a significant difference in SC lengths. But When the means are compared across chrm class, the mouse averages are no longer significant.

How well does SC length predict chromosome class?  Prediction, High rec strains will have more significant p value, given the lower overlap in SC lengths across chrm class. (Note this test can only be done with bivalent level observations, because 1CO or 2CO)

```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE}

```{r Q2.logreg.plots, echo=FALSE}

what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")

```{r SC.Length_t.test, echo=FALSE, eval=FALSE}

#comparing all the strains

T.test_SC.Len_mouse.av <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1],

  Male.poly.Mouse.Table_BivData_4MM$mean_SC[(Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc")&(Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0)]
)#p-value = 0.001567

#1CO
T.test_SC.Len_mouse.av_1CO <- t.test(mouse.avs_4MM$mean.SC_1CO[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean.SC_1CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]

@@

#2CO
T.test_SC.Len_mouse.av_2CO <- t.test(mouse.avs_4MM$mean.SC_2CO[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean.SC_2CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]

#3CO
T.test_SC.Len_mouse.av_3CO <- t.test(mouse.avs_4MM$mean.SC_3CO[mouse.avs_4MM$Rec.group == 1],

  mouse.avs_4MM$mean.SC_3CO[(mouse.avs_4MM$subsp == "Musc") & (mouse.avs_4MM$sex == "male") & (mouse.avs_4MM$Rec.group == 0)]

#Single Biv level
T.test_SC.Len_single.biv <- t.test(Curated_BivData$chromosomeLength[Curated_BivData$Rec.group == 1],
  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)] )

#1CO
T.test_SC.Len_single.biv_1CO <- t.test(
  Curated_BivData$chromosomeLength[( Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 1) ],

      Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0) & (Curated_BivData$hand.foci.count == 1) ] 
  )

#2CO
T.test_SC.Len_single.biv_2CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1) & (Curated_BivData$hand.foci.count == 2)],

                                       Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 2) ] 
  )

  #3CO
  T.test_SC.Len_single.biv_3CO <- t.test(Curated_BivData$chromosomeLength[(Curated_BivData$Rec.group == 1)&(Curated_BivData$hand.foci.count == 3) ],

  Curated_BivData$chromosomeLength[(Curated_BivData$subsp == "Musc")&(Curated_BivData$sex == "male")&(Curated_BivData$Rec.group == 0)&(Curated_BivData$hand.foci.count == 3) ]
  )
```

The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.

```{r long.sc_log.res_meanSC, eval=FALSE}
#Long Biv comparison

#hold off on reporting/ working with the long SC biv resutls

long.SC.mouse.avs_4MM_male <- Long_biv_mouse.avs_4MM[Long_biv_mouse.avs_4MM$sex == "male",]
long.SC.mouse.avs_4MM_Mmale <- long.SC.mouse.avs_4MM_male[long.SC.mouse.avs_4MM_male$subsp == "Musc",]

#divide musc males into groups
long.SC.mouse.avs_4MM_male$Rec.group <- ifelse(grepl("PWD male", long.SC.mouse.avs_4MM_male$category), 1, 
                                    ifelse(grepl("MSM male", long.SC.mouse.avs_4MM_male$category), 1, 0))

long.biv.log.reg <- glm(Rec.group ~ mean_SC, 
                            data=long.SC.mouse.avs_4MM_male, family=binomial(link="logit"))

summary(long.biv.log.reg)#NS effect (likelt underpowered)
```

@@

```{r SC.length.male_tests, echo=FALSE}

@@

```{r SC.length.pairwise.T, echo=FALSE}
#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av

@@

> Is sex a significant effect for SC length? (as predicted)

- Sex is a significant effect for SC length
The results seem to indicate that sex is a significant factor. Consider writing a sub sampling approach (randomize / permute a data set of BivData)

- According to anova, sex effect explains most of the variance in single bivalent SC lengths.

- The Long Biv Data set largely agrees with the full curated dataset

- I caveat I haven't addressed yet, is the XX in the female Biv data averages ---

```{r log.reg_SC.length, echo=FALSE}
#male_mouse.av
#Male.poly.Mouse.Table_BivData_4MM

The mean SC logistic regression model for mouse averages

`r log.reg_SC.length_mouse.av$coefficients`

and for the single bivalent levels

`r log.reg_SC.length_BIV$coefficients`

@@

<!-- 
Main DFs for this section:
Male.poly.Mouse.Table_BivData_4MM
male.bivdata.1CO

>>>>

load(file="~./MLH1repo/data/CleanBivData_19.2.20.RData")

@@

```{r juvimice.list, echo=FALSE}
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )

#REMOVE JUVI mice
#juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
#mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]

@@

#has 
#18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8 is missing the 1CO position -- remove for now
Curated_BivData <- Curated_BivData %>% filter(fileName != "18jul19_31may19_SKIVE_m2_sp1_13.2_rev_8")

@@

#REMOVE JUVI mice
juvi.mouse_CurBiv_table <- mouse.avs_4MM[mouse.avs_4MM$mouse  %in% juvi_mice, ]  
mouse.avs_4MM <- mouse.avs_4MM[! mouse.avs_4MM$mouse  %in% juvi_mice, ]

@@

mouse.avs_4MM$Rec.group.exclu <- ifelse(grepl("PWD male", mouse.avs_4MM$category), 1, 
                                ifelse(grepl("MSM male", mouse.avs_4MM$category), 1, 0))

@@

#unordered for the mixed models (move this to the MM blocks...?)
#keep order for the plots?
male.av_4MM$subsp <- factor(male.av_4MM$subsp, ordered = FALSE)
male.av_4MM$sex <- factor(male.av_4MM$sex, ordered = FALSE)
male.av_4MM$strain <- factor(male.av_4MM$strain, ordered = FALSE)

#REMOVE JUVI mice
juvi.from.male.biv_table <- male.av_4MM[male.av_4MM$mouse  %in% juvi_mice, ]  
mouse.avs_4MM <- male.av_4MM[! male.av_4MM$mouse  %in% juvi_mice, ]

@@

#REMOVE JUVI mice
tot.SC_juvi.male <- totSC.mouse.av.table[totSC.mouse.av.table$mouse  %in% juvi_mice, ]

totSC.mouse.av.table <- totSC.mouse.av.table[! totSC.mouse.av.table$mouse  %in% juvi_mice, ]

@@

# Q2 SC Length

<!-- place set ups for total SC, short biv and long biv here -->

I made a sereies of plots which show that the expected / general pattern of bivalent SC length is positively associated with having more crossovers per chromsomes (possitive correlation of sc length and CO number).  Yet, given the complications from the chromsomes proportions outlined above, comparing the bivalents across chromosome classes is complecated

```{r juvimice.list.copy, echo=FALSE}
juvi_mice <- c(
  "30dec14_WSB_m2","30dec14_WSB_m4","30dec14_WSB_m1","30dec14_WSB_m3","30dec14_WSB_m5",
  "30sep14_G_m1","30sep14_G_m2","30sep14_G_m3","30sep14_G_m4","30sep14_G_m5",
  "12Feb17_LEW_m1","12Feb17_LEW_m2", "13sep17_LEW_m2", "13sep17_LEW_m1","13sep17_LEW_m3",

  "12sep17_PWD_m1","12sep17_PWD_m2","12sep17_PWD_m3",
   "30Dec16_MSM_m2","30Dec16_MSM_m3","30Dec16_MSM_m1"
  )

ppmm <- OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$mouse  %in% juvi_mice, ]   
```

Q2mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                         #mean.first.1CO
                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])
                         )

Q2mouse.av.long.biv.table <- add_strain(Q2mouse.av.long.biv.table)
Q2mouse.av.long.biv.table <- add_subsp(Q2mouse.av.long.biv.table)

Q2mouse.av.long.biv.table$strain <- factor(Q2mouse.av.long.biv.table$strain, ordered = FALSE)
Q2mouse.av.long.biv.table$subsp <- factor(Q2mouse.av.long.biv.table$subsp, ordered = FALSE)

#divide musc males into groups
Q2mouse.av.long.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
                                  ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1,
                                  ifelse(grepl("SKIVE", Q2mouse.av.long.biv.table$strain), 1,
                                           0)))

Q2mouse.av.long.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.long.biv.table$strain), 1, 
                                ifelse(grepl("MSM", Q2mouse.av.long.biv.table$strain), 1, 0))

Q2mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength),

                        #mean.first.1CO
                        norm.first.foci.pos =  mean(PER_Foci_1[which(hand.foci.count == 1)])

Q2mouse.av.short.biv.table <- add_mouse(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table <- add_strain(Q2mouse.av.short.biv.table)
Q2mouse.av.short.biv.table <- add_subsp(Q2mouse.av.short.biv.table)

Q2mouse.av.short.biv.table$strain <- factor(Q2mouse.av.short.biv.table$strain, ordered = FALSE)
Q2mouse.av.short.biv.table$subsp <- factor(Q2mouse.av.short.biv.table$subsp, ordered = FALSE)

Q2mouse.av.short.biv.table$Rec.group <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1,
                                  ifelse(grepl("SKIVE", Q2mouse.av.short.biv.table$strain), 1,
                                           0)))

Q2mouse.av.short.biv.table$Rec.group.exclu <- ifelse(grepl("PWD", Q2mouse.av.short.biv.table$strain), 1, 
                                ifelse(grepl("MSM", Q2mouse.av.short.biv.table$strain), 1, 0))
```

@@

Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)

Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)

Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)

@@

```{r Q2.SC.long.biv_plots, echo=FALSE}

shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")

cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")
mouse.shoo.long.biv
```

```{r Q2.SCshort_plot, echo=FALSE}

```{r Q2_tot.sc_plot, echo=FALSE}

mouse.av_totsc_male <- ggplot(data= totSC.mouse.av.table %>% filter(sex == "male"), aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(), axis.ticks.x=element_blank()) + ylim(c(1200,1850 ))+ggtitle("Mouse av. total SC area")

mouse.av_totsc_male

<!-- basic stats -->

total SC for males (which ar)
X additional cells were added to the male specifi total SC data set discirbed above

-reduced bivalent dataset (long and short) is made up of bined bivalents from X cells and X mice

<!-- tests below;  t.test(high low) OR logistic regression, mixed modlels, glm 
remember the pairwise t.test
Pwise.ttest_Dom.male <- pairwise.t.test(male.Dom.mouse.av$mean_SC, male.Dom.mouse.av$strain)
G-WSB is sig, WSB-LEW are slightly sig

```{r Q2.totSC.ttest, echo=FALSE}

#totalSC high and low rec group

#
Q2tot.sc_t.test_High.Low <- t.test((totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel,

                               (totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "1") )$mean.skel)
#p-value = 0.007

```{r Q2.short.long.ttest, echo=FALSE}

#long and short 
#mouse.av.short.biv.table
#Q2mouse.av.short.biv.table
Q2short.biv_t.test_High.Low <- t.test((Q2mouse.av.short.biv.table %>% filter(Rec.group == "0") )$mean.SC,
                                      (Q2mouse.av.short.biv.table %>% filter(Rec.group == "1") )$mean.SC)

#p-value = 0.9

#mouse.av.long.biv.table
Q2long.biv_t.test_High.Low <- t.test((Q2mouse.av.long.biv.table  %>% filter(Rec.group == "0") )$mean.SC,
                          (Q2mouse.av.long.biv.table  %>% filter(Rec.group == "1") )$mean.SC)
#p-value = 0.2

**we ran t.tests on the SC based metrics to test is the high and low groups are significanly different**

**mouse means of total SC area for Higher rec strains are significantly higher compared to low rec strains (ttest; p = 
`r Q2tot.sc_t.test_High.Low$p.value`).**

**The mouse means for the reduced bivalent datasets are not significantly different between the high and low rec strains (ttest; short p = `r Q2short.biv_t.test_High.Low$p.value` and long p = `r Q2long.biv_t.test_High.Low$p.value`). (the summary of mouse averages for this subset of ) **

this might indicate -there's error in the total SC algorithm

- mouse effects

- low power / sampling bias in reduced data set

- total SC is capturing differences in the XY signal which is excluded from the bivalent dataset

glms

```{r Q2.totSC.glm, echo=FALSE}

#(totSC.mouse.av.table %>% filter(sex == "male") %>% filter(Rec.group == "0") )$mean.skel

Q2.glm_mus.av_totSC_M1 <- glm(mean.skel ~ subsp * strain, data= totSC.mouse.av.table %>% filter(sex == "male"))

sum.Q2.glm_mus.av_totSC_M1  <- summary(Q2.glm_mus.av_totSC_M1)$coefficients

#sum.Q2.glm_mus.av_totSC_M1.coef <-

#sum.Q2.glm_mus.av_totSC_M1$coefficients[,4]   pvalues of all effects
#sum.Q2.glm_mus.av_totSC_M1$coefficients[,1]  coefficients of all effects
#round(sum.Q2.glm_mus.av_totSC_M1$coefficients[,c(1,4)], 6)

Q2.glm_mus.av_totSC_M2 <- glm(mean.skel ~ strain, data= totSC.mouse.av.table %>% filter(sex == "male"))

sum.Q2.glm_mus.av_totSC_M2  <- summary(Q2.glm_mus.av_totSC_M2)
#round(summary(sum.Q2.glm_mus.av_totSC_M2)$coefficients[,c(1,4)], 6)
```

(the subspecis effect is tested in M1)
For M2 (glm including subsp), the two subspecies effects are significant (glm; p= Musculus `r sum.Q2.glm_mus.av_totSC_M1[,4][2]` and Molossinus p= `r sum.Q2.glm_mus.av_totSC_M1[,4][3]`)

strain effects from M1; G, LEW, MSM, SKIVE and TOM

strain effects  (basicall all strain effects are significant) (tech effects / error?)
All strain effects were significant when tested in M2 (Supplemental table X)

For M2 all

```{r Q2.glm.SC.long, echo=FALSE}

#Q2.glm.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
#round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)
#table(mouse.av.long.biv.table$mouse)
#Q2mouse.av.long.biv.table

glmQ2_long.biv_M1 <- glm(mean.SC ~ subsp * strain, data= Q2mouse.av.long.biv.table)

sum.glmQ2_long.biv_M1 <- summary(glmQ2_long.biv_M1)

glmQ2_long.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.long.biv.table)

sum.glmQ2_long.biv_M2 <- summary(glmQ2_long.biv_M2)

#round(sum.glmQ2_long.biv_M2$coefficients[,c(1,4)], 6)
```

**For the long bivalent dataset, In M1 where the subsp are tested, only Musc is significant (glm; Musc p= `r sum.glmQ2_long.biv_M1$coefficients[,4][2]`). The significant strain effects in M1 are slight just MSM (glm; MSM p= `r sum.glmQ2_long.biv_M1$coefficients[,4][7]`).**

**For the long bivalent dataset, In M2 (limited to just the strain effects) a few strain effects are significant (glm; MSM p = `r sum.glmQ2_long.biv_M2$coefficients[,4][5]`, SKIVE `r sum.glmQ2_long.biv_M2$coefficients[,4][7]` and CZECH p = `r sum.glmQ2_long.biv_M2$coefficients[,4][9]`).**

```{r Q2.glm.SC.short, echo=FALSE}
#Q2mouse.av.short.biv.table
#mouse.av.short.biv.table

glmQ2_short.biv_M1 <- glm(mean.SC ~ subsp*strain, data= Q2mouse.av.short.biv.table)
sum.glmQ2_short.biv_M1 <- summary(glmQ2_short.biv_M1)

#round(summary(sum.glmQ2_short.biv_M1)$coefficients[,c(1,4)], 6)

glmQ2_short.biv_M2 <- glm(mean.SC ~ strain, data= Q2mouse.av.short.biv.table)
glmQ2_short.biv_M2 <- summary(glmQ2_short.biv_M2)

In the short bivalent data set -- analyzed by M1, both subspeces are sig (glm; p= Musculus `r sum.glmQ2_short.biv_M1$coefficients[,4][2]` and Molossinus p= `r sum.glmQ2_short.biv_M1$coefficients[,4][3]`). And the strain effec for G is significant (glm; G p =`r sum.glmQ2_short.biv_M1$coefficients[,4][4]`).

For the M2 modeanalyzing the mouse means short bivalents, all strains have significant effect but LEW
G,PWD, MSM, MOLF, SKIVE, KAZ, CZECH.

```{r Q2.SC.long.glm.bivglm, echo=FALSE}

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
#round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
#round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
#round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)

**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.**

@@

@@

@@

@@

@@

what are the pairwise t-test functions -- use pairwise.t.test(Ozone, Month, p.adj = "bonf")
The mouse average and single bivalents are significantly longer in the high rec group p = r T.test_SC.Len_mouse.av$p.value and p=  r T.test_SC.Len_single.biv$p.value.

<!-- junk chunks below -->

```{r Q2.SC.logreg.chrmClass_BivData, echo=FALSE, include=FALSE, eval=FALSE}

```{r Q2.logreg.plots, echo=FALSE, include=FALSE, eval=FALSE}

```{r Q2.SC.Chrm.Class, echo=FALSE, include=FALSE, eval=FALSE}
#REMOVE COMPARES ACROSS FROM FULL BIVDAAT

@@

```{r Q2.SC.log.regression.show.plots, echo=FALSE, include=FALSE}

@@

```{r Q2.SClength.ttest, echo=FALSE, include=FALSE}
#COMPARISONS ACROSS ALL CLASSES

#MOUSE AVE, all
Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

@@

```{r log.reg_SC.length, echo=FALSE, include=FALSE}
#REMOVE THIS COMPARES FROM FULL BIV DATA

#male_mouse.av
#Male.poly.Mouse.Table_BivData_4MM

```{r Q2.SC_class.rec.group_setup, echo=FALSE, include=FALSE}
#REMOVE THIS, IT uses total bivalent data

@@

```{r Q2.SC.Length.stuff, echo=FALSE, include=FALSE}
#REMOVE THIS, COMPARES FULL BIVDATA ACROSS CHRM CLASS

#all pooled SC lengths
#WSB     G   LEW  MOLF   KAZ CZECH 
#    5     9     5     4     3     3

Q1lmer.MM_SC_short_results <- data.frame(sub.sp = Q1lmer.MM_SC_short_subp, 
sex.results =  Q1lmer.MM_SC_short_sex, 
               interaction.pval = Q1lmer.MM_SC_short_interaction,  rand.pval = 
Q1lmer.MM_SC_short_strain)
Q1lmer.MM_SC_short_results
```

```{r Q2.SC.ploy.male, echo=FALSE, include=FALSE}
#REMOVE THIS< COMARES FROM FULL BIVDATA set

# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect

@@

```{r SC.length.male_tests, echo=FALSE, include=FALSE}

@@

```{r SC.length.pairwise.T, echo=FALSE, include=FALSE}

#REMOVE RUNS COMPARISONS ACROSS FULL DATASET
#male.Dom.mouse.av, male.Mol.mouse.av, male.Musc.mouse.av

@@

@@

In order to test /

We tested another aspect of the rec landscape -- focuing on differences between the high and the low rec strains.

To detect if there are significant differences between the high and the low rec strains we performed t.tests across the two groups

(and logistic regressions?)

<!-- set up chunks here
